---
title: "Propensity score matching"
author: "Elizabeth Handorf"
date: "`r Sys.Date()`"
output: 
 html_document:
  toc: true
  toc_float: false
  number_sections: true
---

Example code for running a propensity score-matched analysis.  

This document uses the "colon" dataset in the "survival" package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(plyr)
library(dplyr)
library(arsenal)
library(kableExtra)
library(survival)
library(lubridate)
library(Matching)
library(SuperLearner)
#Change working directory for all code chunks
#knitr::opts_knit$set(root.dir = 'U://')

#Sample data from survival package
dat<-data.frame(colon)
dat<-dat %>% dplyr:::filter(etype==1) %>%
dplyr::filter(rx!="Obs") %>%
dplyr::filter(!is.na(differ) & !is.na(nodes))

```

```{r }

#Set up the dataset for PS estimation

#Indicator for treatment of interest
dat$Lev5FU = as.numeric(dat$rx=="Lev+5FU")

#We will exact match on perforation status, age within 5 years
#All other factors go into the propensity score

#PS estimation via super-learner

#Factor vars
#First make into factors
facts<-dat %>% dplyr::select(differ, extent) 
  facts<- data.frame(lapply(facts, factor))
#Convert factors to numeric
facts_encoded<- model.matrix(~-1+differ+extent,data=facts)
colnames(facts_encoded)<-gsub("\\s", "", colnames(facts_encoded)) 
facts_encoded<-as.data.frame(facts_encoded)
  
#Numeric vars
numeric<- dat %>% dplyr::select(sex, age, obstruct, adhere, nodes, surg)

#Combine the two groups together
X_predictors = cbind(facts_encoded, numeric)

####### Estimate the propensity score
set.seed(147852)
#Run the super-learner with a variety of methods
sl_mod = SuperLearner(Y = dat$Lev5FU, X = X_predictors , family = binomial(),
                      SL.library = c( "SL.glmnet", "SL.randomForest", "SL.bartMachine","SL.ksvm","SL.xgboost"))


pr.Tot<-sl_mod$SL.predict

#May want to save super learner results since this procedure can take a while
#save.image("/SLRes.RData")


```

# Find matches


## Propensity score distribution - pre matched
```{r }

par(mfrow=c(2,1))
hist(pr.Tot[dat$Lev5FU==0],xlim=c(0,1))
hist(pr.Tot[dat$Lev5FU==1],xlim=c(0,1))
```

## Find matches

Matching without replacement. Exact match on perforation and sex, age match within 5 years, propensity score within caliper of 0.01

```{r }

X_match=cbind(pr.Tot,dat$age,dat$perfor, dat$sex)

set.seed(147852)

matches<-Match(Tr=dat$Lev5FU,X=X_match, exact=c(F,F,T,T),M=1,ties=F,replace=F,caliper=c(.01,5,0,0))
summary(matches)

```

## Check match balance

Note: std mean diff = std diff *100  
Rule of thumb: this should be less than 10  
P-values - preferably look at the  Kolmogorov-Smirnov test (ks)  
  
Note: this was originally randomized data.  In some ways, matching is actually making the balance worse!  This is just for example purposes.  

```{r }
#Check the balance

MatchBalance(Lev5FU~as.factor(differ) + as.factor(extent)+sex+ age+ obstruct+ adhere+ nodes+ surg+perfor, data = dat, match.out = matches, ks=TRUE)
```

# Analysis of matched data

See discussion of methods here:
https://cran.r-project.org/web/packages/MatchIt/vignettes/estimating-effects.html#after-pair-matching-without-replacement

```{r, results='asis' }

#Set up the analysis dataset

#Matched cases/controls
#This gives the case and control ides with their respective pair id
pairs<-data.frame(cbind(matches$index.treated, matches$index.control,
                        seq(1:length(matches$index.treated))))
names(pairs)<-c("case","control","pair.id")
pairs_long<-cbind.data.frame(matched_id=c(pairs$case,pairs$control), pair=c(pairs$pair.id,pairs$pair.id))

dat.matched<-dat[pairs_long$matched_id,]
dat.matched$pair.id<-pairs_long$pair

#Now, run analysis
#Survival outcomes
#Re-scale days to months
dat.matched$months<-dat.matched$time/30.4

s<-survfit(Surv(months, status==1)~Lev5FU,data=dat.matched)

#make the table nicer
restab<-summary(s)$table[,c(1,4,7:9)]
rownames(restab)<-c("Lev","Lev+5FU")
kable(restab, digits=c(0,0,1,1,1)) %>% kable_styling()

plot(s, col=c("darkgreen", "navy"), lwd=1.5, main="Overall Survival by treatment type (matched)", xlab="Months")
legend("topright", c("Lev", "Lev+5FU"),col=c("darkgreen","navy"), lty=1)

m1<-coxph(Surv(months, status==1)~Lev5FU,data=dat.matched,
              robust = TRUE, 
              cluster = pair.id)


restab<-coefficients(summary(m1))
restab2<-data.frame(HR=exp(restab[,1]),LCL=exp(restab[,1]-1.96*restab[,3]),UCL=exp(restab[,1]+1.96*restab[,3]),pval=restab[,5])
# More meanigful row names
rownames(restab2)<-c("Lev+5FU vs Lev")

#replace small p-vals
restab2$pval<-round(restab2$pval,3)
restab2$pval[restab2$pval<0.001]<-"<0.001"
#Round the columns to 2 or 3 significant digits
kable(restab2, digits=c(2,2,2,3)) %>% kable_styling()

```

