---
title: "R Markdown Template (without extra notes)"
author: "Elizabeth Handorf"
date: "`r Sys.Date()`"
output: 
 html_document:
  toc: true
  toc_float: false
  number_sections: true
---

This document uses the "colon" dataset in the "survival" package

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


library(plyr)
library(dplyr)
library(arsenal)
library(kableExtra)
library(survival)
library(lubridate)

#Change working directory for all code chunks
#knitr::opts_knit$set(root.dir = 'U://')


```




# Cohort definition

Apply exclusion criteria, create CONSORT table

1. Remove observations (duplicates) with "etype=1"  
2. Remove untreated patients (treatment=observation)  
2. Remove patients>80

```{r, reuslts='asis'}

#empty table to fill in numbers and descriptions
# Number of rows = number of exclusions+1
n.excl<-3
consort.table<-matrix(rep(NA, 2*(n.excl+1)),ncol=2)
i<-1


#Sample data from survival package
#Starting size
dat<-data.frame(colon)

consort.table[i,1]<-"Starting number of observations"
consort.table[i,2]<-dim(dat)[1]
i<-i+1

dat<-dat %>% dplyr:::filter(etype==1)

consort.table[i,1]<-"Remove events not of interest"
consort.table[i,2]<-dim(dat)[1]
i<-i+1

dat<-dat %>% filter(rx!="Obs")

consort.table[i,1]<-"Remove patients who received observation"
consort.table[i,2]<-dim(dat)[1]
i<-i+1

dat<-dat %>% filter(age<=80)

consort.table[i,1]<-"Remove patients over 80"
consort.table[i,2]<-dim(dat)[1]
i<-i+1

kable(consort.table) %>% kable_styling()

```

# Treatment frequency



```{r, reuslts='asis'}
kable(table(dat$rx), col.names = c("Treatment","N")) %>% kable_styling()
```



# Decscribe the dataset - "Table 1"

Cohort characteristics by treatment

```{r, reuslts='asis'}

tab1 <- tableby( rx ~ sex+ as.factor(sex)+age+as.factor(obstruct), data=dat)


#make categorical variables into factors
dat$sex<-factor(dat$sex)

#Customize labels
dat$obstruct<-as.factor(dat$obstruct)
attr(dat$obstruct,'label')  <- 'Obstruction'


tab1 <- tableby( rx ~ sex+ age+obstruct, data=dat, cat.test="fe", numeric.test = "kwt")

kable(summary(tab1, text=TRUE)) %>% kable_styling()
```

# MVA - factors associated with treatment

Logistic regression model

```{r, results="asis"}

#make a binary indicator for treatment
dat$Lev5FU<-as.numeric(dat$rx=="Lev+5FU")

#10 years of age
dat$age10y<-dat$age/10


#Run the logistic regression
m1<-glm(Lev5FU~ age10y+sex+obstruct+perfor+as.factor(extent)+node4
, data=dat, family="binomial")

#Extract the components of interest from the model
restab<-coefficients(summary(m1))

restab2<-data.frame(OR=exp(restab[,1]),LCL=exp(restab[,1]-1.96*restab[,2]),UCL=exp(restab[,1]+1.96*restab[,2]),pval=restab[,4])
                

# More meanigful row names
rownames(restab2)<-c("Intercept",
                    "Age (per 10 year increase)",
                     "Male",
                     "Obstruction",
                     "Perforation",
                     "Spread to muscle (vs mucosa)",
                      "Spread to serosa (vs mucosa)",
                      "Spread to contiguous structures (vs mucosa)",
                     ">4 nodes"
)

#Round the columns to 2 or 3 significant digits
kable(restab2[-1,], digits=c(2,2,2,3)) %>% kable_styling()

```


# Survival Analyses

## Unadjusted

Kaplan-Meier survival curves

```{r, reuslts='asis'}
#Re-scale days to months
dat$months<-dat$time/30.4

s<-survfit(Surv(months, status==1)~Lev5FU, data=dat)

plot(s, col=c("firebrick4","navy"), xlab="Months", ylab="Proportion surviving", main="Progression-free survival")
legend("topright", c("Lev", "Lev+5FU"), col=c("firebrick4","navy"), lty=1)


```

Median survival
```{r, reuslts='asis'}

#make the table nicer
restab<-summary(s)$table[,c(1,4,7:9)]
rownames(restab)<-c("Lev","Lev+5FU")
kable(restab, digits=c(0,0,1,1,1)) %>% kable_styling()


```

Survival at 2 & 5 years

```{r, reuslts='asis'}
#Make a table of survival at specific times

s.T <- summary(s, times=c(24,60))
#extract only the components we want
#convert to a data frame
s.T.mat<-data.frame(Treatment=s.T$strata,
                    time=s.T$time,
                    n.risk=s.T$n.risk, 
                    Surv=s.T$surv, 
                    LCL=s.T$lower, 
                    UCL=s.T$upper)

s.T.mat[,1]<-c("Lev","Lev","Lev+5FU","Lev+5FU")

kable(s.T.mat, digits=c(0,0,0,3,3,3)) %>% kable_styling()
```

## Adjusted Cox PH model


```{r, results="asis"}


m1<-coxph(Surv(months, status==1)~Lev5FU+age10y+sex+obstruct+perfor+as.factor(extent)+node4
, data=dat)

#Format this more nicely
cat(paste("Total observations used = ",m1[["n"]]))

restab<-coefficients(summary(m1))

restab2<-data.frame(HR=exp(restab[,1]),LCL=exp(restab[,1]-1.96*restab[,3]),UCL=exp(restab[,1]+1.96*restab[,3]),pval=restab[,5])

# More meanigful row names
rownames(restab2)<-c("Lev+5FU vs Lev",
                    "Age (per 10 year increase)",
                     "Male",
                     "Obstruction",
                     "Perforation",
                     "Spread to muscle (vs mucosa)",
                      "Spread to serosa (vs mucosa)",
                      "Spread to contiguous structures (vs mucosa)",
                     ">4 nodes"
)

#replace small p-vals
restab2$pval<-round(restab2$pval,3)
restab2$pval[restab2$pval<0.001]<-"<0.001"

#Round the columns to 2 or 3 significant digits
kable(restab2, digits=c(2,2,2,3)) %>% kable_styling()

```